---
title: 'IS606: PROJECT 1'
author: "MUSA T. GANIYU"
date: "February 27, 2016"
output: pdf_document
---

**WHAT IMPACT DOES BODY WEIGHT HAVE ON BRAIN WEIGHT?**

**The data records the average weight of the brain and body for a number of mammal species.**
**There are 62 rows of data.  The 3 data columns include:**
**I,  the index, A1, the brain weight;**
**B,  the body weight.**
**We seek a model of the form:**
**B = A1 * X1.**

**Load data from URL **
```{r}
require(knitr);
weights <- read.table("http://people.sc.fsu.edu/~jburkardt/datasets/regression/x01.txt", skip = 32, header = TRUE, sep = "")
kable(head(weights));
names(weights);
options(warn=-1)
```
** Load require package.**
```{r}
library(plyr);

```
**Note the data columns came with Body & Weight, but its actually "BrainWeight & "BodyWeight" in the data, which needs correction.**
```{r}
names(weights)[names(weights)=="Body"] <- "BrainWeight";
names(weights)[names(weights)=="Weight"] <- "BodyWeight";
str(weights);

names(weights);
kable(head(weights));
```

**Lets checkout its histogram.**
```{r}
hist(weights$BrainWeight);
hist(weights$BodyWeight);
```

**We can deduce that it a Very Rightly Skewed data, an intervention is required.Its getting the Natural logarithm of the data set.**
```{r}

weights_log <- cbind(weights, log(weights$BrainWeight), log(weights$BodyWeight));

kable(head(weights_log));
```

**load required package and attach the data again and plot it in scatterplot 3D.**
```{r}
library(scatterplot3d);

attach(weights_log);
scatterplot3d(BrainWeight, BodyWeight, pch = 20, highlight.3d = TRUE, type = "h", main = "3D ScatterPlots"); 
```


**Removing or seperating new column from initial data set.**
```{r}

weights_log[,c("BrainWeight","BodyWeight")] <- list(NULL);
colnames(weights_log);

a <- plot(weights_log, ylab="Brain Weight",
   plot.type="double", col=1:2, xlab="Body Weight")
legend("topleft", legend=c("Brain Weight","Body Weight"),
  lty=1, col=c(1,2), cex=.8)
abline(a)
```


**Changing the column name.**
```{r}
names(weights_log)[names(weights_log)=="log(weights$BrainWeight)"] <- "BrainWeights";
names(weights_log)[names(weights_log)=="log(weights$BodyWeight)"] <- "BodyWeight";

kable(head(weights_log));
```

**Correlations and covariance shows a strong positive relationship between Brain Weights and Body weight**
```{r}
cor(weights_log, use="complete.obs", method="kendall")


```

**Histogram for Body weight**
```{r}
x <- weights_log$BodyWeight;

hist(x, 
 xlim=c(min(x),max(x)), probability=T, 
   col='purple', xlab='Body Weight', ylab=' Frequency', axes=T,
   main='Natural Logarithm: Multi-modal')
lines(density(x,bw=1), col='red', lwd=2)
```


**Getting the mode of BodyWeight**
```{r}
mode_1 <- table(as.vector(x));
names(mode_1)[mode_1 == max(mode_1)];
```


**Histogram for Brain weight**
```{r}
y <- weights_log$BrainWeight;

hist(y, 
 xlim=c(min(y),max(y)), probability=T, 
   col='purple', xlab='Brain Weight', ylab=' Frequency', axes=T,
   main='Natural Logarithm: Bi-modal')
lines(density(y,bw=1), col='red', lwd=2)
kable(summary(weights_log));
```


**Getting the mode of Brain Weight**
```{r}
mode_2 <- table(as.vector(y));
names(mode_2)[mode_2 == max(mode_2)];

```

**Load required package forecast, for forecasting.**
```{r}
library(forecast);

Acf(weights_log$BodyWeight, lag.max=NULL, type=c("correlation", "partial"), plot=TRUE, main=NULL, xlim=NULL, ylim=NULL, xlab="Lag", ylab=NULL, na.action=na.contiguous); 


taperedacf(weights_log$BodyWeight, lag.max=NULL, type=c("correlation", "partial"), plot=TRUE, calc.ci=TRUE, level=95, nsim=100, xlim=NULL, ylim=NULL, xlab="Lag", ylab=NULL); 


fcast <- forecast(weights_log$BodyWeight, h = ifelse(frequency(weights_log$BodyWeight) > 1, 2 * frequency(weights_log$BodyWeight), 10) , level=c(68, 95, 99.7), fan=FALSE, robust=FALSE, lambda=NULL, find.frequency=FALSE, allow.multiplicative.trend=FALSE);

fcast;

```

**We can conclude based on this analysis that the brain weight actually depend on the Body Weight in Human.**

```{r}
```

**DATA SOURCE: http://people.sc.fsu.edu/~jburkardt/datasets/regression/x01.txt**

**References:**

**http://r.789695.n4.nabble.com/converting-character-to-numeric-td3615259.html**

**http://stackoverflow.com/questions/4605206/drop-data-frame-columns-by-name**

**https://cran.r-project.org/web/packages/forecast/forecast.pdf**

